A comparison of different methods for calculating tangent-stiffness matrices in a massively parallel computational peridynamics code.

### 2014-02-22 response/response.tex

{\color{red}
  \begin{itemize}
    \item 
        We have revised the brief description of peridynamics in page 9 to be more complete:
    \item
        Material from silling2007 is interpreted including: the definition of vector states,
        the derivation of conservation laws in terms of state-based peridynamics, the definition
        of the isotropic elastic solid AKA the Linear elastic Peridynamic Solid.
    \item
        It is the authors' opinion that the additional material mentioned above explains the equations involved
        in the peridynamic models for mechanics, used in relation to the peridynamics code
        studied in the work.

  \end{itemize}}
    \item 
        The manuscript seems to be written in a language more familiar to computer scientists. 
        I would encourage the authors to revise certain parts of the manuscript to be
        presented in a language more familiar to the computational mechanics community.

{\color{red}
  \begin{itemize}
    \item 
        We have revised certain parts of the manuscript for clarity, adding additional information for the reader's
        convenience:
    \item
        In subsection 'Implementing the complex-step method in Peridigm', explanation of what is meant by datatype,
        container class, template, scope and instantiation are included since these concepts may be unfamiliar to
        a computational mechanics audience at least as denoted by these terms. 
    \item 
        In subsection 'The Comparative Study', the way message passing computations can slow runtime is explained
        by an analogy to an experiment given as an exercise for the reader. 
    \item 
        In section 'Conclusions', the term byte-copiable is explained. 
    \item 
        It is the authors' opinion that the above revisions to the language of the manuscript allow the intended audience
        to follow the discussion where concepts from computer science are relevant.
  \end{itemize}}

  \item
     In Section 1.1.1, please include a more comprehensive literature review for
     the CS method, including further references (some of them are listed below).
     In particular, specify what aspects of the CS method have been investigated in
     each reference. Then, you need to clearly mention your contribution within
     the CS method, in contrast to the other references (some comments in this
     regard appear in the manuscript in page 8).

{\color{red}
  \begin{itemize}
     \item 
         The authors have appended a more comprehensive literature review for the CS method to section 1.1.1.
         The aspects of CS investigated by the prior work, as well as any particularly interesting
         developments introduced by those works, have been listed. To argue for the novelty of the instant work,
         each prior work is compared to the instant work in terms of the use of the CS method and the similarity of the fields of research of the works.
     \item
         The literature review concludes with an argument for the novelty of the instant
         work where the prior works are taken as a whole and compared to the instant work. It is the authors' opinion
         that the instant work is novel for at least the reasons set out in the literature review material, where
         the contribution to the CS method is in the comparison to other methods in terms of computational
         time,  parallel implementation, and implementation within \emph{Peridigm} .
  \end{itemize}}

### 2014-02-22 jacobian_calc_CMAME.tex
I.4 response section corrected for grammar and style

\note[Michael]{re: I.4}
It is appropriate to detail the prior work discussing aspects of the
complex-step method in order to indicate what is new in the instant work.  The
CS method was introduced in a work by J. N. Lyness and C. B. Moler
\cite{lyness1967numerical}.  In this prior work, the CS method is identified in
an series form and the computing of analytical derivatives is shown. This prior
work develops the CS method, while the instant work is about an application of
the CS method.

@@

A work by J. C. Newman, W. K. Anderson and L. D. Whitfield, discusses computing
sensitivities in a CFD analysis using complex step and compares the results to
central difference produced results \cite{newman1998}. The work is also notable
for identifying automatic differentiation as a competing method to CS and
suggesting that complex step could be used to compute Jacobians of residuals
for use in Newton-Krylov schemes for solving non-linear systems. The instant
work is novel with respect to the prior work because the instant work
investigates CS as applied to a numerical method for solving non-linear
systems, rather than as in the prior work investigating the derivatives
themselves produced with the CS method.

In a work by A. Perez-Foguet, A. Rodriguez-Ferran, A. Huerta, CS is used to
compute Jacobians for both integrating plastic stress to determine
elastoplastic modulii, but also for computing the Jacobian of the global
problem, both of which are an application of the CS method to the solution of
non-linear systems \cite{perez2000numerical}. These elastoplastic modulii are
used in computing the global Jacobian  and said to be 'algorithmically
consistent' in that they are consistent with the algorithm used to return the
predicted stress to the yield surface. Examples of interest include FEM
simulations on two plastic material models with complex hardening rules.
Comparing the prior work with the instant work, the prior work demonstrates CS
on a serial computer and discusses a classical solid mechanics formulation.
Automatic differentiation is not investigated, and CPU time measurements are not
taken. Scalability studies with a variety of equivalent, progressively more
refined meshes are not done. Scalability as a function of parallel computer
size is not investigated. The instant work is novel with respect to the prior work in
offering a discussion of CS as applied to peridynamics simulation on a parallel
computer, CPU time data, and comparison to automatic differentiation as a competing method.

Another work by A. Perez-Foguet, A. Rodriguez-Ferran, A. Huerta The prior work
applies CS in computing derivatives needed for integrating stress, obtaining
algorithmically consistent elastoplastic modulii, and obtaining the Jacobian of
the global residual, for FEM simulations using the MRS-Lade material model
\cite{perez2012numerical}. In this prior work, CPU runtime is measured.
Comparing the prior work to the instant work, the prior work demonstrates CS on
a serial computer and discusses a classical solid mechanics formulation.
Automatic differentiation is not investigated.  Scalability studies with a
variety of equivalent, progressively more refined meshes are not done.
Scalability as a function of distributed system size is not investigated.  The
instant work is novel with respect to the prior work in offering a discussion
of CS as applied to peridynamics simulation on a parallel computer, and
comparison to automatic differentiation as a competing method.

A work by Joaquim R. R. A. Martins, Ilan M. Kroo and Juan J. Alonso
investigates using CS for computing aerodynamic sensitivities
\cite{martins2000automated}. Additionally higher order accuracy versions of CS
are discussed, as is the task of implementing complex-step in most widely used
general purpose programming languages. Also discussed is the relative ease of
implementing a CS scheme versus an AD based scheme for computing sensitivities
in an established code. The instant work is novel with respect to the prior
work because the instant work investigates CS as applied to a numerical method
for solving non-linear systems, rather than as in the prior work investigating
the using the CS method to obtain sensitivities.

A work by W. K. Anderson, J. C. Newman, D. l. Whitfield and E. J. Nielsen discusses
the CS applied to computing sensitivities in CFD simulations \cite{anderson2001sensitivity}. CS is
compared to AD both in Fortran and C++ applications. In this prior work it is suggested that forward mode
AD is mathematically equivalent to first order CS. The instant work is novel with respect to the prior
work because the instant work investigates CS as applied to a numerical method
for solving non-linear systems, rather than as in the prior work investigating
the derivatives themselves produced with the CS method.

A work by J. Martins, P. Sturdza, J. Alonso demonstrates that the CS method can
be performed by using complex datatypes for the variables of functions
comprising an analysis code. The analysis code can comprise a simulation and
derivatives of response variables called sensitivities can be computed. The
accuracy of this approach is compared to automatic differentiation
\cite{martins2003complex}. The instant work is novel with respect to the prior
work because the instant work investigates CS as applied to a numerical method
for solving non-linear systems, rather than as in the prior work investigating
the derivatives themselves produced with the CS method.

A work by  K. L. Lai and J. L. Crassidis discusses first and second order CS
methods for computing derivatives. Choosing optimal step size
\cite{lai2008extensions}. This prior work discusses computing first and second
order accurate derivatives with CS and choosing optimal step size.  The instant
work discusses using the derivatives computed with the CS method in a numerical
method for solving a non-linear system.

A work by A. H. Al-Mohy and N. J. Higham investigates the use of complex step
in computing Frechet derivatives of matix to matrix functions \cite{al2010complex}. To
compute the residual Jacobian in the peridynamics simulation discussed in the
instant work is analogous because it is computing a Frechet derivative of a
higher dimension function of higher dimension variables. Where the application
of the CS derivatives in the prior work is limited to a particular class of
abstract functions to determine the CS' suitability for use in condition number
estimation, the instant work applies CS to solving a mechanics problem. The
instant work is not about computing Frechet derivatives with CS but moreover using them
in a numerical method to solve a set of equations describing a physical system.

A work by W. Jin, B. H. Dennis, and B. P. Wang discusses using CS to compute
sensitivities in a solid mechanics FEM analysis \cite{jin2010improved}. In the
prior work the derivatives produced by CS themselves are the object of
interest, where the instant work uses CS in a numerical method for solving a
non-linear system.

A work by A. Voorhees, H. Millwater, and R. Bagley, where H. Millwater is an author of
the instant work, discusses the background of the CS method and frames it as a subset
of Fourier Differentiation. CS and Fourier Differentiation are then applied to 
computing sensitivities within a solid mechanics FEA \cite{voorhees2011complex}. A notable
technique demonstrated by this prior work is the representation of complex valued
nodal components as additional degrees of freedom in the FE model, to be interpreted
by a matrix representation of complex numbers during computations, yet without storing
the entire matrix. In the prior work, CS is discussed and related to a broader class of differentiation methods and used
to compute sensitivities. In the instant work, CS is used in a numerical method for solving non-linear systems.

In another work by H. Millwater, C. Quintana, J. Garza and David Wagner the
concept of incorporating complex values into the degrees of freedom of finite
elements is labelled ZFEM \cite{millwater2013application}. Notable in this work
is the discussion of creating user defined element types for use with the
popular \emph{ABAQUS} software suite which enable the use of ZFEM in that
software, \emph{ABAQUS} is a trademark of Dassault Systemes
\cite{systemes2012abaqus}. In this prior work the goal of using CS through the
ZFEM technique is to compute shape sensitivities, while in the instant work CS
is used in a numerical method for solving non-linear systems.

A work by R. Abreu, D. Stich, and J. Morales discusses computing derivatives
with a re-interpreted version of the CS method which features the combined use
of real and imaginary valued step components \cite{abreu2013generalization}.
This prior work suggests that the modified CS method can achieve extended
approximation accuracy up to fourth order.  The prior work discusses using the
CS method to compute derivatives and attempts to extend the method for that
application, where the derivatives themselves are the object of interest. In
contrast, the instant work discusses using the derivatives computed with CS in
a numerical method for solving a non-linear system.

The body of prior literature has introduced the application of CS to
single dimension and multidimensional derivatives, sensitivity analysis,
non-linear solution schemes, the extension of CS for higher accuracy, 
and implementation strategies for CS in new and existing codes.
With respect to the prior work discussing the CS method, the primary novel
aspects of the instant work are then its discussion of the application of CS to
peridynamics simulations run in \emph{Peridigm}, a parallel implementation of
CS and a compilation of CPU time data, where each of CS, AD, FD and CD are
compared.

>>>>

\note[Michael]{re: I.4}
It is appropriate to detail the prior work discussing aspects of the
complex-step method in order to indicate what is new in the instant work.  The
CS method was introduced in a work by J. N. Lyness and C. B. Moler
\cite{lyness1967numerical}.  In this prior work, the CS method is identified in
an infinite series form and the computing of analytical derivatives is shown. This prior
work develops the CS method, while the instant work is about an application of
the CS method.

@@

A work by J. C. Newman, W. K. Anderson and L. D. Whitfield discusses computing
sensitivities in a CFD analysis using complex step and compares the results to
central difference produced results \cite{newman1998}. The work is also notable
for identifying automatic differentiation as a competing method to CS and
suggesting that complex step could be used to compute Jacobians of residuals
for use in Newton-Krylov schemes for solving non-linear systems. The instant
work is novel with respect to the prior work because the instant work
investigates CS as applied to a numerical method for solving non-linear
systems, rather than as in the prior work investigating the derivatives
themselves produced with the CS method.

In a work by A. Perez-Foguet, A. Rodriguez-Ferran, A. Huerta, CS is used to
compute Jacobians for integrating plastic stress to determine
elastoplastic modulii and for computing the Jacobian of the residual for the global
problem, both of which are an application of the CS method to the solution of
non-linear systems \cite{perez2000numerical}. These elastoplastic modulii are
used in computing the global Jacobian  and are said to be 'algorithmically
consistent' in that they are consistent with the algorithm used to return the
predicted stress to the yield surface. Examples of interest include FEM
simulations on two plastic material models with complicated hardening rules.
Comparing the prior work with the instant work, the prior work demonstrates CS
on a serial computer and discusses a classical solid mechanics formulation.
Automatic differentiation is not investigated and CPU time measurements are not
taken. Scalability studies with a variety of equivalent, progressively more
refined meshes are not done. Scalability as a function of parallel computer
size is not investigated. The instant work is novel with respect to the prior work in
offering a discussion of CS as applied to peridynamics simulation on a parallel
computer, CPU time data, and comparison of CS to automatic differentiation as a competing method.

Another work by A. Perez-Foguet, A. Rodriguez-Ferran, A. Huerta applies CS in
computing derivatives needed for integrating stress, obtaining algorithmically
consistent elastoplastic modulii, and obtaining the Jacobian of the global
residual, for FEM simulations using the MRS-Lade material model
\cite{perez2012numerical}. In this prior work, CPU runtime is measured.
Comparing the prior work to the instant work, the prior work demonstrates CS on
a serial computer and discusses a classical solid mechanics formulation.
Automatic differentiation is not investigated.  Scalability studies with a
variety of equivalent, progressively more refined meshes are not done.
Scalability as a function of distributed system size is not investigated.  The
instant work is novel with respect to the prior work in offering a discussion
of CS as applied to peridynamics simulation on a parallel computer, and
comparison of CS to automatic differentiation as a competing method.

A work by Joaquim R. R. A. Martins, Ilan M. Kroo and Juan J. Alonso
investigates using CS for computing aerodynamic sensitivities
\cite{martins2000automated}. Additionally, higher order accuracy versions of CS
are discussed, as is the task of implementing complex-step in most widely used
general purpose programming languages. Also discussed is the relative ease of
implementing a CS scheme versus an AD based scheme for computing sensitivities
in an established code. The instant work is novel with respect to the prior
work because the instant work investigates CS as applied to a numerical method
for solving non-linear systems, rather than as in the prior work investigating
the use of the CS method to obtain sensitivities.

A work by W. K. Anderson, J. C. Newman, D. l. Whitfield and E. J. Nielsen
discusses the CS method as applied to computing sensitivities in CFD simulations
\cite{anderson2001sensitivity}. CS is compared to AD both in Fortran and C++
applications. In this prior work it is suggested that forward mode AD is
mathematically equivalent to first order CS. The instant work is novel with
respect to the prior work because the instant work investigates CS as applied
to a numerical method for solving non-linear systems, rather than as in the
prior work investigating the derivatives themselves produced with the CS
method.

A work by J. Martins, P. Sturdza, J. Alonso demonstrates that the CS method can
be performed by using complex datatypes for the variables of functions
comprising an analysis code \cite{martins2003complex}. The analysis code can
comprise a simulation and derivatives of response variables called
sensitivities can be computed. The accuracy of this approach is compared to
automatic differentiation. The instant work is novel with respect to the prior
work because the instant work investigates CS as applied to a numerical method
for solving non-linear systems, rather than as in the prior work investigating
the derivatives themselves produced with the CS method.

A work by  K. L. Lai and J. L. Crassidis discusses first and second order CS
methods for computing derivatives and choosing optimal step size
\cite{lai2008extensions}. The instant work discusses using the derivatives
computed with the CS method in a numerical method for solving a non-linear
system, while in the prior work the derivatives obtained from the CS method 
are the object of interest.

A work by A. H. Al-Mohy and N. J. Higham investigates the use of complex step
in computing Frechet derivatives of matix to matrix functions
\cite{al2010complex}. To compute the Jacobian of the residual in the
peridynamics simulation as discussed in the instant work is analogous is
computing a Frechet derivative. Where the application of the CS derivatives in
the prior work is limited to a particular class of abstract functions to
determine the CS method's suitability for use in condition number estimation, the
instant work applies CS to solving a solid mechanics problem. The instant work is not
about computing Frechet derivatives with CS but moreover using them in a
numerical method to solve a set of equations describing a physical system.

A work by W. Jin, B. H. Dennis, and B. P. Wang discusses using CS to compute
sensitivities in a solid mechanics FEM analysis \cite{jin2010improved}. In the
prior work the derivatives produced by CS are the object of
interest, whereas the instant work uses CS in a numerical method for solving a
non-linear system.

A work by A. Voorhees, H. Millwater, and R. Bagley discusses the background of
the CS method and frames it as a subset of Fourier Differentiation. CS and
Fourier Differentiation are then applied to computing sensitivities within a
solid mechanics FEA \cite{voorhees2011complex}. A notable technique
demonstrated by this prior work is the representation of complex valued nodal
coordinates with additional degrees of freedom in the FE model, to be
interpreted by a matrix representation of complex numbers during computations.
In the prior work, CS is discussed and related to a broader class of
differentiation methods and used to compute sensitivities. In the instant work,
CS is used in a numerical method for solving non-linear systems.

In another work by H. Millwater, C. Quintana, J. Garza and David Wagner the
concept of incorporating complex values into the degrees of freedom of finite
elements is called ZFEM \cite{millwater2013application}. Notable in this work
is the discussion of creating user defined element types for use with the
popular \emph{ABAQUS} software package which enables the use of ZFEM in that
software. \emph{ABAQUS} is a trademark of Dassault Systemes
\cite{systemes2012abaqus}. In this prior work the goal of using CS through the
ZFEM technique is to compute shape sensitivities, while in the instant work CS
is used in a numerical method for solving non-linear systems.

A work by R. Abreu, D. Stich, and J. Morales discusses computing derivatives
with a version of the CS method which features the combined use
of real and imaginary valued step components \cite{abreu2013generalization}.
This prior work suggests that the modified CS method can achieve extended
approximation accuracy up to fourth order.  The prior work discusses using the
CS method to compute derivatives and attempts to extend the method for that
application, where the derivatives themselves are the object of interest. In
contrast, the instant work discusses using the derivatives computed with CS in
a numerical method for solving a non-linear system.

The body of prior literature has introduced the application of CS to
single dimension and multidimensional derivatives, sensitivity analysis,
non-linear solution schemes, the extension of CS for higher accuracy, 
and implementation strategies for CS in new and existing codes.
With respect to the prior work discussing the CS method, the novel
aspects of the instant work are then its discussion of the application of CS to
peridynamics simulations run in \emph{Peridigm}, a parallel implementation of
CS and a compilation of CPU time data, where each of CS, AD, FD and CD are
compared.
